Effectiveness of community-based marine reserves in small-scale fisheries

### 2018-10-01 docs/paper/Villasenor-Derbez.tex
Incorporated Fios comments, ready for submission?

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Artisanal fisheries face great
challenges since they are difficult to monitor, enforce, and manage.
Combining territorial user rights for fisheries (TURF) with no-take
marine reserves to create TURF-reserves is believed to improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last six
years, 18 TURF-reserves have been implemented in Mexico; their
effectiveness has not been formally evaluated. We combine causal
inference techniques and the Social-Ecological Systems framework to
provide a holistic evaluation of community-based TURF-reserves in three
coastal communities in Mexico. We find that while reserves have not yet
achieved their stated goal of increasing lobster densities, they
continue to receive significant support from the fishing communities. A
lack of ecological and socioeconomic effects likely results from a
combination of factors. First, the lobster fisheries are already well
managed, and it is unlikely that reserves might have a detectable
effect. Second, some of the reserves are not large enough to protect
lobsters' home ranges. Third, some of these reserves might be too young
for the effects to show. However, these reserves have shaped small-scale
fishers' way of thinking about marine conservation, which can provide a
foundation for establishing additional, larger marine reserves needed to
effectively conserve mobile species.

@@

Words in text: 3821 \textbar{} Words in headers: 35 \textbar{} Words
outside text (captions, etc.): 192

@@

\section{Introduction}\label{introduction}

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
artisanal fisheries face great challenges since they tend to be hard to
monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MRs) can greatly improve the performance of coastal
fisheries and the health of the local resouces
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintenance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is
little empirical evidence of their effectiveness and the drivers of
their success \citep{afflerbach_2014-HP,lester_2017}.

The performance of these systems depends on how environmental and social
factors combine and interact. The science of marine reserves has largely
focused on understanding the ecological effects of these areas, which
include increased biomass, species richness, and densities of organisms
within the protected regions, climate change mitigation, and protection
from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
focused on the relationship between socioeconomic and governance
structures and reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective. This is especially important in
social-ecological coastal systems dominated by close interaction and
feedbacks between people and natural resources \citep{ostrom_2009-hg}.

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement, and
reserves can yield benefits similar to systematically-designed reserves
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to
enforce if they are not legally recognized, and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Mexican Civil Society Organizations (CSOs) served as a
link between fishers and government, and created a legal framework that
solves this governance issue. In Mexico, a new norm was created in 2014
allowing fishers to request the legal recognition of community-based
reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero};
\citet{nom}). Fish refuges can be implemented as temporal or partial
reserves, which can protect one, some, or all resources within their
boundaries. Since 2012, 45 Fish Refuges have been created along the
Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18
of them implemented as TURF-reserves. However, their effectiveness has
not yet been formally evaluated and reported in the scientific
literature.

@@

\section{Methods}\label{methods}

\subsection{Study area}\label{study-area}

@@

\subsection{Data collection}\label{data-collection}

@@

\subsection{Data analysis}\label{data-analysis}

@@

We use the SES framework to evaluate each community as a means of
providing an explanation to the biological and socioeconomic results.
The use the SES framework standardizes our analysis and allows us to
communicate our results in a common language across fields. We based our
variable selection primarily on \citet{leslie_2015-na} and
\citet{basurto_2013-oq}, who operationalized and analyzed Mexican
fishing cooperatives using this framework. We also incorporate other
relevant variables known to influence reserve performance following
\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
\ref{table:ses} shows the selected variables, their definition and
selected indicators.

\section{Results}\label{results}

@@

\subsection{Biological effects}\label{biological-effects}

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and fourth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
invertebrate and fish densities (Fig. \ref{fig:indicators}B-D), where
effect sizes oscillated around zero without clear trends. Full tables
with model coefficients are presented in the supplementary materials (S1
Table, S2 Table, S3 Table).

\subsection{Socioeconomic effects}\label{socioeconomic-effects}

@@

\subsection{Governance}\label{governance}

@@

\section{Discussion}\label{discussion}

Our results indicate that these TURF-reserves have not increased lobster
densities. Additionally, no co-benefits were identified when using other
ecological indicators aside from the previously reported buffering
effect that reserves can have to environmental variability in Isla
Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
pertaining landings and revenues showed little to no change after
reserve implementation. The coastal ecosystems where these reserves are
located have been profoundly affected by climatic and oceanographic
extremes, including warming events, extreme storms and prolonged hypoxia
\citep{micheli_2012-EU,woodson}. Despite the lack of evidence of the
effectiveness of these reserves, most of the communities show a positive
perception about their performance and continue to support their
presence \citep{ayer_2018}. Understanding the social-ecological context
in which these communities and their reserves operate might provide
insights as to why this happens.

@@

A first possible explanation for the lack of effectivenes may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
reserves, and our SES analysis suggests that all communities have a
well-established community-based enforcement system. With these
characteristics, one would expect the reserves to be effective. Maria
Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
\textless{} 6 years old) and effects may not yet be evident due to the
short duration of protection, relative to the life histories of the
protected species; other community-based marine reserves in tropical
ecosystems may take up to six years to show a spillover effect
\citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}). Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, fishers
may favor implementation of reserves that pose low fishing costs due to
their small size or location. Our analysis of economic data supports
this, as neither landings nor revenues showed the expected short-term
costs associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2006}. Both lobster fisheries were
certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
Additionally, lobster fisheries are managed via species-specific minimum
catch sizes, seasonal closures, protection of ``berried'' females, and
escapement windows where traps are allowed \citep{dof_website_1993}. It
is uncertain whether such a well-managed fishery will experience
additional benefits from marine reserves. Additionally,
\citet{gelcich_2008} have shown that TURFs alone can have greater
biomass and richness than areas operating under open acces. This might
reduce the difference between indicators from the TURF and reserve
sites, making it difficult to detect such a small change. Further
research should focus on evaluating sites in the reserve, TURF, and open
access areas or similar Fish Refuges established without the presence of
TURFs where the impact of ther eserves might be larger.

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat, like coral reefs. These sites can serve
as an insurance against uncertainty and errors in fisheries management,
as well as environmental shocks
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU,aalto}. Self-regulation
of fishing effort (\emph{i.e.} reduction in harvest) can serve as a way
to compensate for future declines associated to environmental variation
\citep{finkbeiner_2018}. Embarking in a marine conservation project can
bring the community together, which promotes social cohesion and builds
social capital \citep{fulton_2019}. Showing commitment to marine
conservation and sustainable fishing practices allows fishers to have
greater bargaining power and leverage over fisheries management
\citep{prezramrez_2012}. Furthermore, the lack of effectiveness observed
in these reserves may not be generalizable to other reserves established
under the same legal frameowrk (\emph{i.e.} Fish Refges) in Mexico, and
future research should aim at evaluating other areas that have also been
established as bottom-up processes but without the presence of TURFs
(\emph{e.g.} \citet{dof_websiteC_2012}), or others established through a
top-down process (\emph{i.e.} \citet{dof_websiteU_2018}).

Community-based marine reserves in small-scale fisheries can be helpful
conservation and fishery management tools when appropriately
implemented. Lessons learned from these cases can guide implementation
of community-based marine reserves elsewhere. For the particular case of
the marine reserves that we evaluate, the possibility of expanding
reserves or merging existing polygons into larger areas should be
evaluated and proposed to the communities. At the broader scale, having
full community support surely represents an advantage, but it is
important that marine reserves meet essential design principles such as
size and placement. Community-based marine reserves might have more
benefits that result from indirect effects of the reserves, which should
be taken into account when evaluating the outcomes of similar projects.

@@

\bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}

@@

\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l}
\hline
Indicator & Units\\
\hline
\multicolumn{2}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
\hline
\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}Income from target species & M MXP\\
\hline
\hspace{1em}Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

>>>>

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial user rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last six
years, 18 TURF-reserves have been implemented in Mexico following a new
regulation; their effectiveness has not been formally evaluated. We
combine causal inference techniques and the Social-Ecological Systems
framework to provide a holistic evaluation of community-based
TURF-reserves in three coastal communities in Mexico. We find that while
reserves have not yet achieved their stated goal of increasing the
density of lobster and other benthic invertebrates, they continue to
receive significant support from the fishing communities. A lack of
clear ecological and socioeconomic effects likely results from a
combination of factors. First, lobster and other invertebrate fisheries
are already well managed, and it is unlikely that reserves might have a
detectable effect. Second, some of the reserves are not large enough to
protect mobile species, like lobster. Third, some of these reserves
might be too young for the effects to show. Fourth, variable and extreme
oceanographic conditions have impacted harvested populations. However,
these reserves have shaped small-scale fishers' way of thinking about
marine conservation, which can provide a foundation for establishing
additional, larger marine reserves needed to effectively conserve mobile
species.

@@

Words in text: 3718 \textbar{} Words in headers: 35 \textbar{} Words
outside text (captions, etc.): 192

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
small-scale fisheries face great challenges since they tend to be hard
to monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MRs) can greatly improve the performance of coastal
fisheries and the health of the local resources
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintenance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is
little empirical evidence of their effectiveness and the drivers of
their success \citep{afflerbach_2014-HP,lester_2017}.

The performance of these systems depends on how environmental and social
factors combine and interact. The science of marine reserves has largely
focused on understanding the ecological effects of these areas, which
include increased biomass, species richness, and densities of organisms
within the protected regions, climate change mitigation, and protection
from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
focused on the relationship between socioeconomic and governance
structures and reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective. This is especially important in
social-ecological coastal systems dominated by close interactions and
feedbacks between people and natural resources \citep{ostrom_2009-hg}.

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement, and
reserves can yield benefits similar to systematically-designed reserves
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to
enforce if they are not legally recognized, and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Mexican Civil Society Organizations (CSOs) served as a
link between fishers and government, and created a legal framework that
solves this governance issue. In Mexico, a new norm was created in 2014
allowing fishers to request the legal recognition of community-based
reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero};
\citet{nom}). Fish refuges can be implemented as temporal or partial
reserves, which can protect one, some, or all resources within their
boundaries. Since 2006, 45 Fish Refuges have been created along the
Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18
of them implemented as TURF-reserves legally recognized as ``Fish
Refuges''. However, their effectiveness has not yet been formally
evaluated and reported in the scientific literature.

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{study-area}{%
\subsection{Study area}\label{study-area}}

@@

\hypertarget{data-collection}{%
\subsection{Data collection}\label{data-collection}}

@@

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

@@

We use the SES framework to evaluate each community. The use the SES
framework standardizes our analysis and allows us to communicate our
results in a common language across fields by using a set of previously
defined variables and indicators. We based our variable selection
primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
operationalized and analyzed Mexican fishing cooperatives using this
framework. We also incorporate other relevant variables known to
influence reserve performance following \citet{difranco_2016-Xw} and
\citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
variables, their definition and selected indicators.

\hypertarget{results}{%
\section{Results}\label{results}}

@@

\hypertarget{biological-effects}{%
\subsection{Biological effects}\label{biological-effects}}

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and fourth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no significant changes were detected in fish
biomass or invertebrate and fish densities (Fig.
\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Full tables with model coefficients are presented
in the supplementary materials (S1 Table, S2 Table, S3 Table).

\hypertarget{socioeconomic-effects}{%
\subsection{Socioeconomic effects}\label{socioeconomic-effects}}

@@

\hypertarget{governance}{%
\subsection{Governance}\label{governance}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Our results indicate that these TURF-reserves have not increased lobster
densities. Additionally, no co-benefits were identified when using other
ecological indicators aside from the previously reported buffering
effect that reserves can have to environmental variability in Isla
Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
pertaining landings and revenues showed little to no change after
reserve implementation. Despite the lack of evidence of the
effectiveness of these reserves, most of the communities show a positive
perception about their performance and continue to support their
presence \citep{ayer_2018}. Understanding the social-ecological context
in which these communities and their reserves operate might provide
insights as to why this happens.

@@

A first possible explanation for the lack of effectiveness may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
reserves, and our SES analysis suggests that all communities have a
well-established community-based enforcement system. With these
characteristics, one would expect the reserves to be effective. Maria
Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
\textless{} 6 years old) and effects may not yet be evident due to the
short duration of protection, relative to the life histories of the
protected species; other community-based marine reserves in tropical
ecosystems may take up to six years to show a spillover effect
\citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}). Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, fishers
may favor implementation of reserves that pose low fishing costs due to
their small size or location. Our analysis of economic data supports
this hypothesis, as neither landings nor revenues showed the expected
short-term costs associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
fisheries were certified by the Marine Stewardship Council
\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
via species-specific minimum catch sizes, seasonal closures, protection
of ``berried'' females, and escapement windows where traps are allowed
\citep{dof_website_1993}. It is uncertain whether such a well-managed
fishery will experience additional benefits from marine reserves.
Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
greater biomass and richness than areas operating under open access.
This might reduce the difference between indicators from the TURF and
reserve sites, making it difficult to detect such a small change.
Further research should focus on evaluating sites in the reserve, TURF,
and open access areas or similar Fish Refuges established without the
presence of TURFs where the impact of the reserves might be larger.

Finally, extreme conditions, including prolonged hypoxia, heat waves,
and storms have affected both the Pacific and Caribbean regions, with
large negative impacts of coastal marine species and ecosystems
\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
where these reserves are located have been profoundly affected by these
events \citep{micheli_2012-EU,woodson}. Effects of protection might be
eliminated by the mortalities associated with these extreme conditions.

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat. These sites can serve as an insurance
against uncertainty and errors in fisheries management, as well as mild
environmental shocks
\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
serve as a way to compensate for future declines associated to
environmental variation \citep{finkbeiner_2018}. Embarking in a marine
conservation project can bring the community together, which promotes
social cohesion and builds social capital \citep{fulton_2019}. Showing
commitment to marine conservation and sustainable fishing practices
allows fishers to have greater bargaining power and leverage over
fisheries management \citep{prezramrez_2012}. Furthermore, the lack of
effectiveness observed in these reserves should not be generalizable to
other reserves established under the same legal framework (\emph{i.e.}
Fish Refuges) in Mexico, and future research should aim at evaluating
other areas that have also been established as bottom-up processes but
without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
or others established through a top-down process (\emph{i.e.}
\citet{dof_websiteU_2018}).

Community-based marine reserves in small-scale fisheries can be helpful
conservation and fishery management tools when appropriately
implemented. Lessons learned from these cases can guide implementation
of community-based marine reserves elsewhere. For the particular case of
the marine reserves that we evaluate, the possibility of expanding
reserves or merging existing polygons into larger areas should be
evaluated and proposed to the communities. At the broader scale, having
full community support surely represents an advantage, but it is
important that marine reserves meet essential design principles such as
size and placement. Community-based marine reserves might have more
benefits that result from indirect effects of the reserves, particularly
providing resilience to shocks and management errors, and promoting
social cohesion, which should be taken into account when evaluating the
outcomes of TURF-reserves.

@@

\bibliographystyle{frontiersinSCNS_ENG_HUMS}
\bibliography{references}

@@

\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l}
\hline
Indicator & Units\\
\hline
\multicolumn{2}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}Income from target species & M MXP\\
\hline
\hspace{1em}Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}
